---
title: "Edgeweights figure"
author: "D. G. Gannon"
date: "January 2021"
output:
  github_document:
    pandoc_args: --webtex
bibliography: /Users/dusty/Documents/zotero_library.bib
---


```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

  require(rstan)
  require(tidyverse)
  require(here)
  require(ggridges)
  require(raster)
  require(gridExtra)
  require(latex2exp)
  
# user-defined functions
  source(here(
    "R", "prediction_functions.R"
  ))

```


### Load data

```{r}

fit <- readRDS(here("Data","mfit_full_dist_interactions_wm_wc_re100.rds"))
load(here("Data", "Indiv_lndscp_gen.RData"))
forest <- raster(here("Data", "forest.tif"))

```

### Extract posteriors of parameters and predict new $\bf W$

```{r}

# extract parameter posteriors
  beta_post <- as.data.frame(rstan::extract(fit, pars="beta"))
  alpha_post <- as.data.frame(rstan::extract(fit, pars="alpha_raw"))
  tau_post <- as.data.frame(rstan::extract(fit, pars="tau"))
  

```

### List of covariates

* Intercept

* geographic distance between plants $i$ and $j$ (Km)

* mean plant density within five meters around $i$ and $j$

* mean canopy cover above $i$ and $j$

* mean proportion of forest within a 500m buffer around $i$ and $j$

* distance $\times$ plant density interaction

* distance $\times$ canopy cover interaction

* distance $\times$ isolation (proportion of forest within 500m)

* $\mathbb{I}$($i$ and $j$ from the same meadow complex)

* $mathbb{I}$($i$ and $j$ from the same meadow)

### Plot posteriors for regression coefficients

```{r}

 labels <- expression(
      "Geographic distance (km)",
      "Mean plant density",
      "Mean canopy cover",
      "Mean isolation",
      "Distance"%*%"density",
      "Distance"%*%"canopy cover",
      "Distance"%*%"isolation",
      bold(I)*" {"*italic("i,j")*" belong to same complex}",
      bold(I)*" {"*italic("i,j")*" belong to same meadow}"
    )

  post_estims <- data.frame(
    expl_var = rep(c(
      "Intercept",
      "geodist",
      "dens",
      "cover",
      "isolation",
      "dist_dens",
      "dist_cover",
      "dist_isol",
      "same_complex",
      "same_meadow"
    ), each=nrow(beta_post)),
    beta = as.double(
      unlist(beta_post)
    )
  )

# remove the intercept for plotting
  post_estims <- post_estims[-which(post_estims$expl_var == "Intercept"), ]

# convert to factor for the sake of ggplot
  post_estims <- post_estims %>%
    mutate(
      f.expl_var = factor(
        expl_var,
        levels = unique(expl_var)
      )
    )
  
# reverse levels for the plot
  
  post_estims <- post_estims %>%
    mutate(
      f.expl_var_rev = fct_rev(f.expl_var)
    )
  

  
# create plot
  fill_col <- "#ccccccff"
  
 effects_plot <-  ggplot(data = post_estims, aes(x=beta, y = f.expl_var_rev))+
    geom_density_ridges(scale=2, size=0.2, bandwidth=0.05,
                        fill=fill_col, alpha=0.7)+
    geom_vline(xintercept = 0, size=0.2, linetype="dashed",
               alpha=0.5) +
    theme(panel.background = element_blank(),
          panel.grid.major.x = element_line(colour = "grey", size = 0.1),
          axis.text.y = element_text(colour = "black", size=12),
          axis.title = element_text(size=14))+
    xlab(expression(beta[j]))+
    scale_y_discrete(name="Posterior Density",
                     breaks=waiver(),
                     labels=labels[length(labels):1])
 
 svg(filename = here("Figures", "lndscp_gen_effs.svg"),
     height = 6, width = 6, family = "arial")
   effects_plot
 dev.off()

```

### Create data to predict over

Create hypothetical plants for prediction located at the center of the sampled locations for each meadow.

```{r}

# find node centers
  node_data <- group_by(samp_dat_sub, MEADOW_ID) %>% 
                  summarise(., 
                    n=n(), 
                    node_x=mean(X_COORD), 
                    node_y=mean(Y_COORD),
                    plant_density=mean(PL_IN_5M), 
                    mean_cover=mean(COVER)/100,
                    meadow_area=mean(MEADOW_AREA),
                    forest_100m=mean(PROP_FOREST_100M_PL)
                  )

# add complex info
  mdw_complex <- unique(samp_dat_sub[,c("MEADOW_ID", "COMPLEX")])
  node_data$complex <- mdw_complex$COMPLEX

# compute the X array for the new plants
  N_new <- nrow(node_data)
  P <- dim(beta_post)[2]
  X_new <- array(data=1, dim = c(N_new, N_new, P))
  
# distance

  X_new[,,2] <- sqrt(
    dist_matrix(
      cbind(node_data$node_x, node_data$node_y)
    )
  )/1000

# mean plant density
 X_new[,,3] <- sym_matrix(node_data$plant_density)
 
# mean canopy cover
 X_new[,,4] <- sym_matrix(node_data$mean_cover)
 
# mean isolation
 X_new[,,5] <- sym_matrix(node_data$forest_100m)
 
# distance by plant density
 X_new[,,6] <- X_new[,,2]*X_new[,,3]
 
# distance by cover
 X_new[,,7] <- X_new[,,2]*X_new[,,4]
 
# distance by isolation
 X_new[,,8] <- X_new[,,2]*X_new[,,5]
 
# within meadow
 for(i in 1:N_new){
   for(j in 1:N_new){
     X_new[,,9][i,j] <- as.numeric(node_data$complex[i] == node_data$complex[j])
   }
 }
 
# within complex
 X_new[,,10] <- diag(x=1, nrow = N_new, ncol=N_new)
 
 dimnames(X_new)[[3]] <- c("intercept",
                             "dist",
                             "avg_pl_density_ij", 
                             "avg_cover_ij",
                             "avg_forest100_ij",
                             "dens_by_dist",
                             "cover_by_dist",
                             "forest_by_dist",
                             "Isame_complex",
                             "Isame_meadow")

```


### Create $\bf Z$ matrix

```{r}

Z_new <- diag(x=1, nrow = N_new, ncol = N_new)

```


### Plot an estimate of the gene flow network

```{r}

# get posterior means for beta and alpha

  beta_post_mean <- as.double(apply(beta_post, 2, mean))
  alpha_post_mean <- apply(alpha_post, 2, mean)
  tau_post_mean <- mean(tau_post$tau)

  
# make the weights matrix
  W <- make_W(N_new, X_new, 
              beta = beta_post_mean,
              Z=Z_new,
              alpha = alpha_post_mean,
              tau = tau_post_mean)
  
# convert to a pairwise dataset
  lines_df <- tibble(
    meadow_i = rep(node_data$MEADOW_ID, N_new),
    meadow_j = rep(node_data$MEADOW_ID, each=N_new),
    weight = as.vector(W, mode = "double")
  )
  
# remove ii connections
  lines_df <- lines_df[
    -which(lines_df$meadow_i == lines_df$meadow_j),
    ]
  
# merge in location data
  
  lines_df_sp <- merge(lines_df, 
                       node_data[,c("MEADOW_ID", "node_x", "node_y")],
                       by.x="meadow_i",
                       by.y="MEADOW_ID")
  lines_df_sp <- merge(lines_df_sp,
                       node_data[,c("MEADOW_ID", "node_x", "node_y")],
                       by.x="meadow_j",
                       by.y="MEADOW_ID")
  
  names(lines_df_sp)[4:length(names(lines_df_sp))] <- c(
    "x_i", "y_i", "x_j", "y_j"
  )
  
# merge in complex information
  lines_df_sp <- merge(
    lines_df_sp,
    node_data[,c("MEADOW_ID", "complex")],
    by.x="meadow_j",
    by.y="MEADOW_ID"
  )
  names(lines_df_sp)[which(names(lines_df_sp)=="complex")] <- "complex_j"
  
  lines_df_sp <- merge(
    lines_df_sp,
    node_data[,c("MEADOW_ID", "complex")],
    by.x="meadow_i",
    by.y="MEADOW_ID"
  )
  names(lines_df_sp)[which(names(lines_df_sp)=="complex")] <- "complex_i"
  
# re_sort
  lines_df_sp <-
    lines_df_sp[order(lines_df_sp$meadow_i, lines_df_sp$meadow_j), ]
  
# create complex-specific datasets
  cm_df <- lines_df_sp[
    which(
      lines_df_sp$complex_i == "CM" &
      lines_df_sp$complex_j == "CM"
    ),
  ]
  
  m2_df <- lines_df_sp[
    which(
      lines_df_sp$complex_i == "M2" &
      lines_df_sp$complex_j == "M2"
    ),
  ]
  
  m1_df <- lines_df_sp[
    which(
      lines_df_sp$complex_i == "M1" &
      lines_df_sp$complex_j == "M1"
    ),
  ] 
  
  lom_df <- lines_df_sp[
    which(
      lines_df_sp$complex_i == "LOM" &
      lines_df_sp$complex_j == "LOM"
    ),
  ]   
  
# # scale the weight by the midrange
  stretch_factor <-5
  
  lines_df_sp <- lines_df_sp %>%
    mutate(
      weight_scl = weight/min(weight)-1
    )
  lines_df_sp$weight_scl <- lines_df_sp$weight_scl/max(lines_df_sp$weight_scl)*2

  
 # CM 
  cm_df <- cm_df %>%
    mutate(
      weight_scl = exp(stretch_factor*(log(weight) - min(log(weight))))-1
    )
  cm_df$weight_scl <- cm_df$weight_scl/max(cm_df$weight_scl)*2
  
 # M2
  m2_df <- m2_df %>%
    mutate(
      weight_scl = exp(stretch_factor*(log(weight) - min(log(weight))))-1
    )
  m2_df$weight_scl <- m2_df$weight_scl/max(m2_df$weight_scl)*2
  
 # M1
  m1_df <- m1_df %>%
    mutate(
      weight_scl = exp(stretch_factor*(log(weight) - min(log(weight))))-1
    )
  m1_df$weight_scl <- m1_df$weight_scl/max(m1_df$weight_scl)*2
  
# LOM
  lom_df <- lom_df %>%
    mutate(
      weight_scl = exp(stretch_factor*(log(weight) - min(log(weight))))-1
    )
  lom_df$weight_scl <- lom_df$weight_scl/max(lom_df$weight_scl)*2
  
# create raster df
  forest_pts <- rasterToPoints(forest, spatial = T)
  forest_df <- data.frame(forest_pts)
  rm(forest_pts)
  
# plot colors
  forest_col <- "#ccccccff"
  mdw_col <- "white"
  point_col <- "#d40000ff"
  line_col <- "#00ccffff"
  
  map_theme <- theme(
    panel.background = element_blank(),
    legend.position = "none",
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    plot.title = element_text(family = "Arial", size = 18)
  )
  
  cm <- ggplot()+
    geom_raster(data = forest_df, aes(x=x, y=y, fill=forest))+
    geom_segment(data = cm_df,
                 aes(x=x_i, y=y_i, xend=x_j, yend=y_j),
                 size=cm_df$weight_scl,
                 color=line_col,
                 alpha=cm_df$weight_scl/max(cm_df$weight_scl))+
    geom_point(data = node_data,
               aes(x=node_x, y=node_y), color=point_col,
               size=3)+
    geom_segment(aes(
      x= 567200, 
      y = 4904300,
      xend = 567700,
      yend = 4904300,
    ), color="black", size=2) + 
    scale_fill_gradient(low = mdw_col, high = forest_col)+
    xlim(c(567000, 569000))+
    ylim(c(4902100, 4904500))+
    ggtitle("Carpenter Mountain")+
    map_theme
  
  m2 <- ggplot()+
    geom_raster(data = forest_df, aes(x=x, y=y, fill=forest))+
    geom_segment(data = m2_df,
                 aes(x=x_i, y=y_i, xend=x_j, yend=y_j),
                 size=m2_df$weight_scl,
                 color=line_col,
                 alpha=m2_df$weight_scl/max(m2_df$weight_scl))+
    geom_point(data = node_data,
               aes(x=node_x, y=node_y), color=point_col,
               size=3)+
    geom_segment(
      aes(x=568200, y=4900300, xend=568700, yend=4900300),
      color="black",
      size=2
    ) +
    scale_fill_gradient(low = mdw_col, high = forest_col)+
    xlim(c(568000, 570800))+
    ylim(c(4898900, 4900500))+
    ggtitle("M2")+
    map_theme
  
  m1 <- ggplot()+
    geom_raster(data = forest_df, aes(x=x, y=y, fill=forest))+
    geom_segment(data = m1_df,
                 aes(x=x_i, y=y_i, xend=x_j, yend=y_j),
                 size=m1_df$weight_scl,
                 color=line_col,
                 alpha=m1_df$weight_scl/max(m1_df$weight_scl))+
    geom_point(data = node_data,
               aes(x=node_x, y=node_y), color=point_col,
               size=3)+
    geom_segment(
      aes(x=570700, y=4898900, xend=571200, yend=4898900),
      color="black",
      size=2
    ) +
    scale_fill_gradient(low = mdw_col, high = forest_col)+
    xlim(c(570500, 572000))+
    ylim(c(4898000, 4899000))+
    ggtitle("M1")+
    map_theme
  
  lom <- ggplot()+
    geom_raster(data = forest_df, aes(x=x, y=y, fill=forest))+
    geom_segment(data = lom_df,
                 aes(x=x_i, y=y_i, xend=x_j, yend=y_j),
                 size=lom_df$weight_scl,
                 color=line_col,
                 alpha = lom_df$weight_scl/max(lom_df$weight_scl))+
    geom_point(data = node_data,
               aes(x=node_x, y=node_y), color=point_col,
               size=3)+
    geom_segment(
      aes(x=568700, y=4895550, xend=569200, yend=4895550),
      color="black",
      size=2
    ) +
    scale_fill_gradient(low = mdw_col, high = forest_col)+
    xlim(c(568500, 572000))+
    ylim(c(4894500, 4895750))+
    ggtitle("Lookout Mountain")+
    map_theme
    
  
  all <- ggplot()+
    geom_raster(data = forest_df, aes(x=x, y=y, fill=forest))+
    geom_segment(data = lines_df_sp,
                 aes(x=x_i, y=y_i, xend=x_j, yend=y_j),
                 size=lines_df_sp$weight_scl,
                 color=line_col,
                 alpha=lines_df_sp$weight_scl/max(lines_df_sp$weight_scl))+
    geom_point(data = node_data,
               aes(x=node_x, y=node_y), color=point_col,
               size=3)+
    geom_segment(
      aes(x=571800, y=4904000, xend=572800, yend=4904000),
      color="black",
      size=2.5
    ) +
    scale_fill_gradient(low = mdw_col, high = forest_col)+
    ggtitle("Frizzel Ridge") +
    map_theme +
    theme(
      plot.title = element_text(size = 24)
    )
  
  plot_lo <- rbind(
    c(1,1,1,2,2),
    c(1,1,1,2,2),
    c(1,1,1,3,3),
    c(1,1,1,3,3),
    c(1,1,1,4,4),
    c(1,1,1,4,4),
    c(1,1,1,5,5),
    c(1,1,1,5,5)
  )
  
png(filename = here("Figures", "genetic_networks.png"),
    height = 3600, width = 3300, units = "px", res=300)  
  grid.arrange(all, cm, m2, m1, lom,
             layout_matrix=plot_lo)
dev.off()
```




