---
title: "Edgeweights figure"
author: "D. G. Gannon"
date: "January 2021"
output:
  github_document:
    pandoc_args: --webtex
bibliography: /Users/dusty/Documents/zotero_library.bib
---


```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

  require(rstan)
  require(CholWishart)
  require(tidyverse)
  require(here)
  require(ggridges)
  require(raster)
  require(extrafont)
  require(gridExtra)

# functions
 #compute distance matrix among observations in matrix m
  dist_matrix <- function(m){
    n <- nrow(m)
    ones <- rep(1,n)
    Gmat <- m%*%t(m) 
    diag_G <- diag(Gmat)
    return(ones%*%t(diag_G) -2*Gmat + diag_G%*%t(ones))
  }
  
# construct the weights matrix given array X,
#  vector beta, matrix Z, and vector alpha
  
  make_W <- function(size, X, beta, Z, alpha, tau){
    W_upper <- matrix(data=0, nrow = size, ncol = size)
    for(i in 1:size){
      for(j in i:size){
        W_upper[i,j] <- exp(X[i,j,]%*%beta + Z[i,]%*%alpha*tau)
      }
    }
    W_upper <- `diag<-`(W_upper, 0)
    return(W_upper + t(W_upper))
  }
  
# create symmetric matrices with {i,j} pair means
  
  sym_matrix <- function(v){
    m <- matrix(data=0, nrow = length(v), ncol = length(v))
    for (i in 1:length(v)) {
      for(j in i:length(v)){
        m[i,j] <- 0.5*(v[i] + v[j])
      }
    }
    m_zdiag <- m
    m_zdiag <- `diag<-`(m,0)
    return(m + t(m_zdiag))
  }

```


### Load data

```{r}

fit <- readRDS(here("Data","mfit_full_dist_interactions_wm_wc_re.rds"))
load(here("Data", "Indiv_lndscp_gen.RData"))
forest <- raster(here("Data", "forest.tif"))

```

### Extract posteriors of parameters and predict new $\bf W$

```{r}

# extract parameter posteriors
  beta_post <- as.data.frame(rstan::extract(fit, pars="beta"))
  alpha_post <- as.data.frame(rstan::extract(fit, pars="alpha_raw"))
  tau_post <- as.data.frame(rstan::extract(fit, pars="tau"))
  

```

### List of covariates

* Intercept

* geographic distance between plants $i$ and $j$ (Km)

* mean plant density within five meters around $i$ and $j$

* mean canopy cover above $i$ and $j$

* mean proportion of forest within a 500m buffer around $i$ and $j$

* distance $\times$ plant density interaction

* distance $\times$ canopy cover interaction

* distance $\times$ isolation (proportion of forest within 500m)

* $\mathbb{I}$($i$ and $j$ from the same meadow complex)

* $mathbb{I}$($i$ and $j$ from the same meadow)

### Plot posteriors for regression coefficients

```{r}

  post_estims <- data.frame(
    expl_var = rep(c(
      "Intercept",
      "Geographic distance (Km)",
      "Mean plant density",
      "Mean canopy cover",
      "Mean isolation",
      "Distance*plant density",
      "Distance*canopy cover",
      "Distance*isolation",
      "I{same complex}",
      "I{same meadow}"
    ), each=nrow(beta_post)),
    beta = as.double(
      unlist(beta_post)
    )
  )

# remove the intercept for plotting
  post_estims <- post_estims[-which(post_estims$expl_var == "Intercept"), ]

# convert to factor for the sake of ggplot
  post_estims <- post_estims %>%
    mutate(
      f.expl_var = factor(expl_var,
                          levels = c(
      "Geographic distance (Km)",
      "Mean plant density",
      "Mean canopy cover",
      "Mean isolation",
      "Distance*plant density",
      "Distance*canopy cover",
      "Distance*isolation",
      "I{same complex}",
      "I{same meadow}" 
                          ))
    )
  
# reverse levels for the plot
  
  post_estims <- post_estims %>%
    mutate(
      f.expl_var_rev = fct_rev(f.expl_var)
    )
  

  
# create plot
  fill_col <- "#003049"
  
 effects_plot <-  ggplot(data = post_estims, aes(x=beta, y = f.expl_var_rev))+
    geom_density_ridges(scale=2, size=0.2, bandwidth=0.05,
                        fill=fill_col, alpha=0.7)+
    geom_vline(xintercept = 0, size=0.2, linetype="dashed",
               alpha=0.5) +
    theme(panel.background = element_blank(),
          panel.grid.major.x = element_line(colour = "grey", size = 0.1),
          axis.text.y = element_text(colour = "black", size=12, family = "Times New Roman"),
          axis.title = element_text(size=14, family = "Times New Roman"))+
    ylab("Posterior density")+
    xlab(expression(beta[j]))
 
 png(filename = here("Figures", "lndscp_gen_effs.png"),
     height = 1800, width = 1800, units = "px", res = 300)
   effects_plot
 dev.off()

```

### Create data to predict over

Create hypothetical plants for prediction located at the center of the sampled locations for each meadow.

```{r}

# find node centers
  node_data <- group_by(samp_dat_sub, MEADOW_ID) %>% 
                  summarise(., 
                    n=n(), 
                    node_x=mean(X_COORD), 
                    node_y=mean(Y_COORD),
                    plant_density=mean(PL_IN_5M), 
                    mean_cover=mean(COVER)/100,
                    meadow_area=mean(MEADOW_AREA),
                    forest_500m=mean(PROP_FOREST_500M_PL)
                  )

# add complex info
  mdw_complex <- unique(samp_dat_sub[,c("MEADOW_ID", "COMPLEX")])
  node_data$complex <- mdw_complex$COMPLEX

# compute the X array for the new plants
  N_new <- nrow(node_data)
  P <- dim(beta_post)[2]
  X_new <- array(data=1, dim = c(N_new, N_new, P))
  
# distance

  X_new[,,2] <- sqrt(
    dist_matrix(
      cbind(node_data$node_x, node_data$node_y)
    )
  )/1000

# mean plant density
 X_new[,,3] <- sym_matrix(node_data$plant_density)
 
# mean canopy cover
 X_new[,,4] <- sym_matrix(node_data$mean_cover)
 
# mean isolation
 X_new[,,5] <- sym_matrix(node_data$forest_500m)
 
# distance by plant density
 X_new[,,6] <- X_new[,,2]*X_new[,,3]
 
# distance by cover
 X_new[,,7] <- X_new[,,2]*X_new[,,4]
 
# distance by isolation
 X_new[,,8] <- X_new[,,2]*X_new[,,5]
 
# within meadow
 for(i in 1:N_new){
   for(j in 1:N_new){
     X_new[,,9][i,j] <- as.numeric(node_data$complex[i] == node_data$complex[j])
   }
 }
 
# within complex
 X_new[,,10] <- diag(x=1, nrow = N_new, ncol=N_new)
 
 dimnames(X_new)[[3]] <- c("intercept",
                             "dist",
                             "avg_pl_density_ij", 
                             "avg_cover_ij",
                             "avg_forest500_ij",
                             "dens_by_dist",
                             "cover_by_dist",
                             "forest_by_dist",
                             "Isame_complex",
                             "Isame_meadow")

```


### Create $\bf Z$ matrix

```{r}

Z_new <- diag(x=1, nrow = N_new, ncol = N_new)

```


### Plot an estimate of the gene flow network

```{r}

# get posterior means for beta and alpha

  beta_post_mean <- as.double(apply(beta_post, 2, mean))
  alpha_post_mean <- apply(alpha_post, 2, mean)
  tau_post_mean <- mean(tau_post$tau)

  
# make the weights matrix
  W <- make_W(N_new, X_new, 
              beta = beta_post_mean,
              Z=Z_new,
              alpha = alpha_post_mean,
              tau = tau_post_mean)
  
# convert to a pairwise dataset
  lines_df <- tibble(
    meadow_i = rep(node_data$MEADOW_ID, N_new),
    meadow_j = rep(node_data$MEADOW_ID, each=N_new),
    weight = as.vector(W, mode = "double")
  )
  
# remove ii connections
  lines_df <- lines_df[
    -which(lines_df$meadow_i == lines_df$meadow_j),
    ]
  
# merge in location data
  
  lines_df_sp <- merge(lines_df, 
                       node_data[,c("MEADOW_ID", "node_x", "node_y")],
                       by.x="meadow_i",
                       by.y="MEADOW_ID")
  lines_df_sp <- merge(lines_df_sp,
                       node_data[,c("MEADOW_ID", "node_x", "node_y")],
                       by.x="meadow_j",
                       by.y="MEADOW_ID")
  
  names(lines_df_sp)[4:length(names(lines_df_sp))] <- c(
    "x_i", "y_i", "x_j", "y_j"
  )
  
# merge in complex information
  lines_df_sp <- merge(
    lines_df_sp,
    node_data[,c("MEADOW_ID", "complex")],
    by.x="meadow_j",
    by.y="MEADOW_ID"
  )
  names(lines_df_sp)[which(names(lines_df_sp)=="complex")] <- "complex_j"
  
  lines_df_sp <- merge(
    lines_df_sp,
    node_data[,c("MEADOW_ID", "complex")],
    by.x="meadow_i",
    by.y="MEADOW_ID"
  )
  names(lines_df_sp)[which(names(lines_df_sp)=="complex")] <- "complex_i"
  
# re_sort
  lines_df_sp <-
    lines_df_sp[order(lines_df_sp$meadow_i, lines_df_sp$meadow_j), ]
  
# create complex-specific datasets
  cm_df <- lines_df_sp[
    which(
      lines_df_sp$complex_i == "CM" &
      lines_df_sp$complex_j == "CM"
    ),
  ]
  
  m2_df <- lines_df_sp[
    which(
      lines_df_sp$complex_i == "M2" &
      lines_df_sp$complex_j == "M2"
    ),
  ]
  
  m1_df <- lines_df_sp[
    which(
      lines_df_sp$complex_i == "M1" &
      lines_df_sp$complex_j == "M1"
    ),
  ] 
  
  lom_df <- lines_df_sp[
    which(
      lines_df_sp$complex_i == "LOM" &
      lines_df_sp$complex_j == "LOM"
    ),
  ]   
  
# # scale the weight by the midrange
  stretch_factor <-5
  
  lines_df_sp <- lines_df_sp %>%
    mutate(
      weight_scl = exp(1.5*(log(weight) - min(log(weight))))
    )
  lines_df_sp$weight_scl <- lines_df_sp$weight_scl/max(lines_df_sp$weight_scl)*2

  
 # CM 
  cm_df <- cm_df %>%
    mutate(
      weight_scl = exp(stretch_factor*(log(weight) - min(log(weight))))
    )
  cm_df$weight_scl <- cm_df$weight_scl/max(cm_df$weight_scl)*2
  
 # M2
  m2_df <- m2_df %>%
    mutate(
      weight_scl = exp(stretch_factor*(log(weight) - min(log(weight))))
    )
  m2_df$weight_scl <- m2_df$weight_scl/max(m2_df$weight_scl)*2
  
 # M1
  m1_df <- m1_df %>%
    mutate(
      weight_scl = exp(stretch_factor*(log(weight) - min(log(weight))))
    )
  m1_df$weight_scl <- m1_df$weight_scl/max(m1_df$weight_scl)*2
  
# LOM
  lom_df <- lom_df %>%
    mutate(
      weight_scl = exp(stretch_factor*(log(weight) - min(log(weight))))
    )
  lom_df$weight_scl <- lom_df$weight_scl/max(lom_df$weight_scl)*2
  
# create raster df
  forest_pts <- rasterToPoints(forest, spatial = T)
  forest_df <- data.frame(forest_pts)
  rm(forest_pts)
  
# plot colors
  forest_col <- "grey"
  mdw_col <- "white"
  point_col <- "#D62828"
  line_col <- "#003049"
  
  map_theme <- theme(
    panel.background = element_blank(),
    legend.position = "none",
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    plot.title = element_text(family = "Times New Roman", size = 18)
  )
  
  cm <- ggplot()+
    geom_raster(data = forest_df, aes(x=x, y=y, fill=forest))+
    geom_segment(data = cm_df,
                 aes(x=x_i, y=y_i, xend=x_j, yend=y_j),
                 size=cm_df$weight_scl,
                 color=line_col,
                 alpha=cm_df$weight_scl/max(cm_df$weight_scl))+
    geom_point(data = node_data,
               aes(x=node_x, y=node_y), color=point_col,
               size=3)+
    geom_segment(aes(
      x= 567200, 
      y = 4904300,
      xend = 567700,
      yend = 4904300,
    ), color="black", size=2) + 
    scale_fill_gradient(low = mdw_col, high = forest_col)+
    xlim(c(567000, 569000))+
    ylim(c(4902100, 4904500))+
    ggtitle("Carpenter Mountain")+
    map_theme
  
  m2 <- ggplot()+
    geom_raster(data = forest_df, aes(x=x, y=y, fill=forest))+
    geom_segment(data = m2_df,
                 aes(x=x_i, y=y_i, xend=x_j, yend=y_j),
                 size=m2_df$weight_scl,
                 color=line_col,
                 alpha=m2_df$weight_scl/max(m2_df$weight_scl))+
    geom_point(data = node_data,
               aes(x=node_x, y=node_y), color=point_col,
               size=3)+
    geom_segment(
      aes(x=568200, y=4900300, xend=568700, yend=4900300),
      color="black",
      size=2
    ) +
    scale_fill_gradient(low = mdw_col, high = forest_col)+
    xlim(c(568000, 570800))+
    ylim(c(4898900, 4900500))+
    ggtitle("M2")+
    map_theme
  
  m1 <- ggplot()+
    geom_raster(data = forest_df, aes(x=x, y=y, fill=forest))+
    geom_segment(data = m1_df,
                 aes(x=x_i, y=y_i, xend=x_j, yend=y_j),
                 size=m1_df$weight_scl,
                 color=line_col,
                 alpha=m1_df$weight_scl/max(m1_df$weight_scl))+
    geom_point(data = node_data,
               aes(x=node_x, y=node_y), color=point_col,
               size=3)+
    geom_segment(
      aes(x=570700, y=4898900, xend=571200, yend=4898900),
      color="black",
      size=2
    ) +
    scale_fill_gradient(low = mdw_col, high = forest_col)+
    xlim(c(570500, 572000))+
    ylim(c(4898000, 4899000))+
    ggtitle("M1")+
    map_theme
  
  lom <- ggplot()+
    geom_raster(data = forest_df, aes(x=x, y=y, fill=forest))+
    geom_segment(data = lom_df,
                 aes(x=x_i, y=y_i, xend=x_j, yend=y_j),
                 size=lom_df$weight_scl,
                 color=line_col,
                 alpha = lom_df$weight_scl/max(lom_df$weight_scl))+
    geom_point(data = node_data,
               aes(x=node_x, y=node_y), color=point_col,
               size=3)+
    geom_segment(
      aes(x=568700, y=4895550, xend=569200, yend=4895550),
      color="black",
      size=2
    ) +
    scale_fill_gradient(low = mdw_col, high = forest_col)+
    xlim(c(568500, 572000))+
    ylim(c(4894500, 4895750))+
    ggtitle("Lookout Mountain")+
    map_theme
    
  
  all <- ggplot()+
    geom_raster(data = forest_df, aes(x=x, y=y, fill=forest))+
    geom_segment(data = lines_df_sp,
                 aes(x=x_i, y=y_i, xend=x_j, yend=y_j),
                 size=lines_df_sp$weight_scl,
                 color=line_col,
                 alpha=lines_df_sp$weight_scl/max(lines_df_sp$weight_scl))+
    geom_point(data = node_data,
               aes(x=node_x, y=node_y), color=point_col,
               size=3)+
    geom_segment(
      aes(x=571800, y=4904000, xend=572800, yend=4904000),
      color="black",
      size=2.5
    ) +
    scale_fill_gradient(low = mdw_col, high = forest_col)+
    ggtitle("Frizzel Ridge") +
    map_theme +
    theme(
      plot.title = element_text(size = 24)
    )
  
  plot_lo <- rbind(
    c(1,1,1,2,2),
    c(1,1,1,2,2),
    c(1,1,1,3,3),
    c(1,1,1,3,3),
    c(1,1,1,4,4),
    c(1,1,1,4,4),
    c(1,1,1,5,5),
    c(1,1,1,5,5)
  )
  
png(filename = here("Figures", "genetic_networks.png"),
    height = 3600, width = 3300, units = "px", res=300)  
  grid.arrange(all, cm, m2, m1, lom,
             layout_matrix=plot_lo)
dev.off()
```


```{r}

# 3D figure

X_comb <- array(dim = c(dim(X)[c(1,2)],9))
  X_comb[,,c(1:2)] <- X[,,c(1:2)]
# now average them
  X_comb[,,3] <- (X[,,"plant_density_i"] +
    X[,,"plant_density_j"])/2
  
  X_comb[,,4] <- (X[,,"cover_i"]+
                   X[,,"cover_j"])/2
  
  X_comb[,,5] <- (X[,,"forest_500m_i"]+
                   X[,,"forest_500m_j"])/2
  X_comb[,,6] <- X_comb[,,2]*X_comb[,,4]
  X_comb[,,7] <- X_comb[,,2]*X_comb[,,5]
  X_comb[,,8] <- X[,,10]
  X_comb[,,9] <- X[,,11]
  
  dimnames(X_comb)[[3]] <- c("intercept", "geo_dist_ij",
                             "avg_pl_density_ij", "avg_cover_ij",
                             "avg_forest500_ij", "dist_by_cover",
                             "dist_by_connect", "I_meadow", "I_complex")

pw_data_ij$scatter <- col_alscatter[lower.tri(col_alscatter)]
pw_data_ij$mean_iso <- X_comb[,,5][lower.tri(X_comb[,,5])]

png(filename = here("Figures","isol_dist_3D1.png"), height = 1800,
    width = 1800, res=300)
with(pw_data_ij,{
     scatter3D(x=mean_iso, y=dist_ij, z=scatter,
               phi = 35, theta = 55,
               xlab="Patch isolation", ylab = "Distance",
               zlab="allelic cov.")
}
)
dev.off()


png(filename = here("Figures","isol_dist_3D2.png"), height = 1800,
    width = 1800, res=300)
with(pw_data_ij,{
     scatter3D(x=mean_iso, y=dist_ij, z=scatter,
               phi = 20, theta = 0,
               xlab="Patch isolation", ylab = "Distance",
               zlab="allelic cov.")
}
)
dev.off()


png(filename = here("Figures","isol_dist_3D3.png"), height = 1800,
    width = 1800, res=300)
with(pw_data_ij,{
     scatter3D(x=mean_iso, y=dist_ij, z=scatter,
               phi = 20, theta = 90,
               xlab="Patch isolation", ylab = "Distance",
               zlab="allelic cov.")
}
)
dev.off()

```


