---
title: "Population differentiation at missense mutations in flower genes"
author: "D. G. Gannon"
date: "March 2021"
output:
  github_document:
    pandoc_args: --webtex
bibliography: /Users/dusty/Documents/zotero_library.bib
---

```{r setup, include=FALSE, warning=FALSE}

knitr::opts_chunk$set(echo = TRUE)

  require(tidyverse)
  require(here)
  require(vcfR)

  source(file = here(
     "R",
     "amova_functions.R"
  ))
```

## Load data

```{r}

 ves <- read.table(file = here("Data", "AQFO_vepreds.txt"), header = T)

```

load names of flower genes

```{r}

fgenes <- read.table(file = 
              here("Data", "flower_genes.txt"),
              header = T,
              sep = "\t"
          )

```


Remove "version" identifiers from gene IDs in the `ensemble_vep` output.

```{r}

ves <- ves %>% mutate(
  gene2 = str_replace_all(ves$Gene,
                pattern = ".v3.1",
                replacement = "")
)

```

Match variants to flower genes

```{r}

flower_variants <- ves[
  which(ves$gene2 %in% fgenes$gene_id),
]

```

Remove duplicates and full variant effect dataset

```{r}
  flower_variants <- flower_variants[-which(duplicated(flower_variants$Location)), ]

  rm(ves)
```

Filter out snps not in the "coding" dataset

```{r}

  snps <- read.vcfR(file = here(
     "Data",
     "AQFO_snps_cd.vcf"
  ))

  flower_variants <- flower_variants[
     which(flower_variants$Uploaded_variation %in% snps@fix[,"ID"]),
  ]
  
# summarize the flower-related genes
  flower_variants %>%
    group_by(Consequence) %>%
    summarise(
      n=n()
    )
 

```



## Create weights matrix with which to compute distances

A weights matrix can be included in AMOVA to account for ancillary information on the markers being using to compute distances between pairs of individuals. We (arbitrarily) weight differences at loci where the alternative allele results in a missense mutation with 1, those that result in a synonymous mutation with 0.25, and any loci upstream or downstream of genes as 0.5. We then weight the off-diagonals elements $w_{ij}, i\ne j$ as the product of the diagonal elements and the inverse of the chromosome distance $d_{ij}$ (in kilobases) between the two loci such that

$$
w_{ij} = \frac{1}{d_{ij}}(w_{ii}w_{jj})\cdot \mathbb{I}\{i,j \text{ on the same chromosome} \},
$$

where $\mathbb{I}\{\cdot\}$ denotes the indicator function returning 1 if the argument in brackets in satisfied and 0 otherwise.


```{r}

# define weights of the diagonal based on effect of variants
  flower_variants$weight <- 0.5
  flower_variants$weight[
     which(flower_variants$Consequence == "missense_variant")
  ] <- 1
  flower_variants$weight[
     which(flower_variants$Consequence == "synonymous_variant")
  ] <- 0.25

  
# construct weights matrix
  W <- diag(x=flower_variants$weight)

  
```

## Compute distances

```{r}

# get dosage genotypes of individuals
 dosage <- allele_load(snps)
 dosage_fgenes <- 
    as.data.frame(t(dosage[ which(snps@fix[,"ID"] %in% flower_variants$Uploaded_variation), ]))
 dosage_fgenes$Sample <- rownames(dosage_fgenes)
 
# order rows by pop, then by group (isolated vs. connected)
 samp_dat <- read.table(
    file = here(
       "Data",
       "sequenced_samples.txt"
    ),
    header = T,
    sep = "\t"
 )
## create group column
 samp_dat$iso <- substr(samp_dat$STRATUM, 3,3)
 
## subset for merge
 samp_dat_sub <- samp_dat[,which(
    names(samp_dat) %in% c("Sample", "MEADOW_ID", "iso")
 )]
 
## merge the two datasets
 dosage_fgenes <- merge(
    samp_dat_sub,
    dosage_fgenes,
    by = "Sample"
 )
 
## order by group then population
 dosage_fgenes <- 
    dosage_fgenes[order(dosage_fgenes$iso, dosage_fgenes$MEADOW_ID), ]

# compute stat of interest
 grp_diff_test <- perm_group_test(dosage_fgenes, W=W, nperm = 1000)
 
# Observed ratio (Connected/Isolated)
 obs_diff <- grp_diff_test$obs["C"] - grp_diff_test$obs["I"]
 
 perm_diffs <- as.double(grp_diff_test$permutations[,1] - grp_diff_test$permutations[,2])
 
# permutational p-value
 mean(perm_diffs < obs_diff) + mean(perm_diffs > -obs_diff)
 
# permutational CIs

## Connected meadows
 quantile(
   grp_diff_test$permutations[,1],
   probs = c(0.025,0.975)
 )
 
## Isolated meadows
 quantile(
   grp_diff_test$permutations[,2],
   probs = c(0.025,0.975)
 )


```




